---
title: "Affective Control Training: Analyses from Protocol"
author: "Jovita T. Leung"
date: "28/03/2021"
output: 
  html_document: 
  toc: yes
  toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup
```{r message=F, warning=F}
rm(list=ls())
```


```{r message=F, warning=F}
# Load libraries
library(dplyr)
library(doBy)
library(lmerTest) 
library(Hmisc)
library(anchors)
library(car) # Anova
library(emmeans)
library(ggfortify)
library(data.table)
library(psych)
library(ggeffects) # extract model predictions
library(sjPlot) # make LMM summary tables
library(jtools) # summarise LMM models
library(ggsignif)
library(ggpubr) # arrange plots
library(Bolstad) # bayesian t test
library(reshape2)
library(lavaan) #SEM
```


```{r}
# Import data
demo=read.csv("~/Dropbox/PhD/04_phd-studies/Affective Control Training Study/ac_training/processed_data/demographics_group.csv")
nback=read.csv("~/Dropbox/PhD/04_phd-studies/Affective Control Training Study/ac_training/processed_data/nback.csv")
train=read.csv("~/Dropbox/PhD/04_phd-studies/Affective Control Training Study/ac_training/processed_data/training_info.csv")
all=read.csv("~/Dropbox/PhD/04_phd-studies/Affective Control Training Study/ac_training/data/281019wodupl2.csv") # full data from Susanne
```


```{r warning=F, message=F}
demo$X=NULL
demo[c("subjno","gender","ses","Group")]=lapply(demo[c("subjno","gender","ses","Group")],factor)
demo$Group=factor(demo$Group, levels = c("0","1"),
                  labels=c("control","AffeCT"))

train$X=NULL
train[c("subjno","no_of_sessions_under_10_mins","no_of_sessions_over_10_mins","no_of_moodtracker_only_sessions",
        "total_number_of_days_1_session_only","total_number_of_days_2plus_sessions","days_betw_T1T2")] = lapply(
          train[c("subjno","no_of_sessions_under_10_mins","no_of_sessions_over_10_mins","no_of_moodtracker_only_sessions",
                  "total_number_of_days_1_session_only","total_number_of_days_2plus_sessions","days_betw_T1T2")], 
          factor)

nback$X=NULL
nback$subjno=as.factor(nback$subjno)
nback[2:37]=lapply(nback[2:37],as.numeric)
```

```{r}
nback1=merge(demo,nback, by="subjno")
```


## Training groups comparison
Following the protocol, two training groups will be compared for potential confounds before hypothesis testing. Any variables showing significant group differences at baseline will be added to subsequent group comparison as covariate. 

```{r}
# Age
bayes.t.test(decimal_age ~ Group, demo) #no significant group difference (p=.202)
```

```{r}
# Gender
chisq.test(demo$gender,demo$Group) #no significant group difference (p=.616)
```

```{r}
# Parental education
plot(demo$ses)
chisq.test(demo$ses, demo$Group) #noz significant group difference (p=.902)
```

```{r}
# IQ
plot(demo$iq_sum)
bayes.t.test(iq_sum ~ Group, demo)
summary(glm(demo$iq_sum~demo$Group)) #no significant group difference (p=.135)
```

```{r}
plot(demo$test_days)
bayes.t.test(test_days ~ Group, demo)
summary(glm(demo$test_days~demo$Group)) #no significant group difference (p=.360)
```

## H1: Affective control training hypothesis
Affective control can be improved in adolescents. To investigate this hypothesis, we will compare individuals’ performances on the affective n-back task across the two training groups.

```{r}
# H1: nback - d prime 
nback2=subset(nback1, select=c("subjno","Group","decimal_age","Scores.dPrimeEmo_T4_1","Scores.dPrimeEmo_T4_2"))
nback_long<-reshape2::melt(nback2, id = c("Group", "subjno","decimal_age"),
                                 measured = c("Scores.dPrimeEmo_T4_1",
                                              "Scores.dPrimeEmo_T4_2"))

nback_long$time=factor(nback_long$variable, levels=c("Scores.dPrimeEmo_T4_1","Scores.dPrimeEmo_T4_2"),
                       labels = c("pre","post"))
names(nback_long)[names(nback_long)=="value"]="dPrimeEmo"
nback_long$variable=NULL
nback_long=na.omit(nback_long)
summary(nback_long)

# mean d' by group and time
summaryBy(dPrimeEmo ~ Group + time, FUN=mean, na.rm=TRUE, data=nback_long)
```
```{r}
# H1: dPrimeEmo model
m.1= lmer(dPrimeEmo ~ Group*time + (1|subjno), 
          data=nback_long)
Anova(m.1, type=3)
summary(m.1) # nothing significant
```

```{r}
# H1a: nback - RT
nback3=subset(nback1, select=c("subjno","Group","decimal_age","Scores.meanRTCorrHitsEmo_T4_1","Scores.meanRTCorrHitsEmo_T4_2"))
nback_long_RT<-reshape2::melt(nback3, id = c("Group", "subjno","decimal_age"),
                                 measured = c("Scores.meanRTCorrHitsEmo_T4_1",
                                              "Scores.meanRTCorrHitsEmo_T4_2"))

nback_long_RT$time=factor(nback_long_RT$variable, levels=c("Scores.meanRTCorrHitsEmo_T4_1","Scores.meanRTCorrHitsEmo_T4_2"),
                       labels = c("pre","post"))
names(nback_long_RT)[names(nback_long_RT)=="value"]="meanRT_emo"
nback_long_RT$variable=NULL
nback_long_RT=na.omit(nback_long_RT)
summary(nback_long_RT)

# mean RT by group and time
summaryBy(meanRT_emo ~ Group + time, FUN=mean, na.rm=TRUE, data=nback_long_RT)
```

```{r}
# H1a: meanRT_emo model
m.1a= lmer(meanRT_emo ~ Group*time + (1|subjno), 
          data=nback_long_RT)
Anova(m.1a, type=3)
summary(m.1a)
```
```{r}
# H1a: post-hoc
emmeans(m.1a, pairwise~time, adjust="Bonferroni") #meanRT is significantly lower post-training
emmeans(m.1a, pairwise~time|Group, adjust="Bonferroni") #meanRT for both groups are lower post-training
emmeans(m.1a, pairwise~Group|time, adjust="Bonferroni") #AffeCT showed lower RT than control post-training. No difference pre-training
```
Significant main effects of Time but not Group. Significant interactions between Group and Time. Mean RT is lower for both groups post-training - this is probably due to familiarity with the task so not very interesting. 


*** 
## H2: Affective control facets hypothesis
AC training will improve inhibition, updating and shifting. 

### Inhibition
```{r}
# Import data
inhib=read.csv("~/Dropbox/PhD/04_phd-studies/Affective Control Training Study/ac_training/processed_data/stroop.csv")
inhib$X=NULL
inhib$subjno=as.factor(inhib$subjno)

inhib1=merge(demo,inhib, by="subjno")
```

```{r}
# H2: Inhibition - accuracy (sad)
inhib2=subset(inhib1, select = c("subjno","Group","Scores.accSadIncongRaw_T3_1","Scores.accSadIncongRaw_T3_2", 
                                 "Scores.accSadCongRaw_T3_1","Scores.accSadCongRaw_T3_2"))
inhib2$sad_diff_1=inhib2$Scores.accSadIncongRaw_T3_1 - inhib2$Scores.accSadCongRaw_T3_1 #calculate difference between congruent and incongruent trials
inhib2$sad_diff_2=inhib2$Scores.accSadIncongRaw_T3_2 - inhib2$Scores.accSadCongRaw_T3_2
inhib2=na.omit(inhib2)
inhib3=subset(inhib2, select = c("subjno","Group","sad_diff_1","sad_diff_2"))

inhib_long_acc=melt(inhib3, id = c("Group","subjno"),
                     measured = c("sad_diff_1","sad_diff_2"))

inhib_long_acc$time=factor(inhib_long_acc$variable, levels=c("sad_diff_1","sad_diff_2"),
                       labels = c("pre","post"))
names(inhib_long_acc)[names(inhib_long_acc)=="value"]="acc_sad_diff"
inhib_long_acc$variable=NULL
summary(inhib_long_acc)

# mean accuracy by group and time
summaryBy(acc_sad_diff ~ Group + time, FUN= mean, na.rm=TRUE, data=inhib_long_acc)
```

```{r}
# H2: acc_sad_incong model 
m.2=lmer(acc_sad_diff ~ Group*time + (1|subjno), inhib_long_acc)
summary(m.2) # Nothing significant
```
```{r}
# plot for H2
mod=data.frame(emmeans(m.2, pairwise ~ time|Group,adjust="Bonferroni", type="response")$emmeans)
ggplot(inhib_long_acc, aes(Group, acc_sad_diff, color=time, fill=time))+geom_violin(alpha=0.3)+
  geom_jitter(position=position_jitterdodge(dodge.width = 0.9), size=1, alpha=0.1)+
  geom_errorbar(data=mod, aes(Group, emmean, ymin = lower.CL, ymax = upper.CL),width = 0.05,size  = 0.5, position = position_dodge(width=0.9))+
  theme_bw()
```


```{r}
# H2a: Inhibition - RT (sad)
inhib4=subset(inhib1, select = c("subjno","Group","Scores.RTSadIncong_T3_1","Scores.RTSadIncong_T3_2",
                                 "Scores.RTSadCong_T3_1","Scores.RTSadCong_T3_2"))
inhib4$RT_diff_1=inhib4$Scores.RTSadIncong_T3_1 - inhib4$Scores.RTSadCong_T3_1
inhib4$RT_diff_2=inhib4$Scores.RTSadIncong_T3_2 - inhib4$Scores.RTSadCong_T3_2
inhib5=subset(inhib4, select = c("subjno","Group","RT_diff_1","RT_diff_2"))
inhib5=na.omit(inhib5)

inhib_long_RT=melt(inhib5, id = c("Group","subjno"),
                     measured = c("RT_diff_1","RT_diff_2"))

inhib_long_RT$time=factor(inhib_long_RT$variable, levels=c("RT_diff_1","RT_diff_2"),
                       labels = c("pre","post"))
names(inhib_long_RT)[names(inhib_long_RT)=="value"]="RT_sad_diff"
inhib_long_RT$variable=NULL
summary(inhib_long_RT)

# mean RT by group and time
summaryBy(RT_sad_diff ~ Group + time, FUN= mean, na.rm=TRUE, data=inhib_long_RT)
```

```{r}
# H2a: RT_sad_incong model
m.2a=lmer(RT_sad_diff ~ Group*time + (1|subjno), inhib_long_RT)
summary(m.2a)
```

```{r}
# H2a: plot
mod=data.frame(emmeans(m.2a, pairwise ~ time|Group, adjust="Bonferroni", type="response")$emmeans)
ggplot(inhib_long_RT, aes(Group, RT_sad_diff, color=time, fill=time))+geom_violin(alpha=0.3)+
  geom_jitter(position=position_jitterdodge(dodge.width = 0.9), size=1, alpha=0.1)+
  geom_errorbar(data=mod, aes(Group, emmean, ymin = lower.CL, ymax = upper.CL),width = 0.05,size  = 0.5, position = position_dodge(width=0.9))+
  theme_bw()
```


### Updating
```{r}
# Import data
update=read.csv("~/Dropbox/PhD/04_phd-studies/Affective Control Training Study/ac_training/processed_data/digit_span.csv")
update$X=NULL
update$subjno=as.factor(update$subjno)

update1=merge(demo,update, by="subjno")
```

```{r}
# H2: updating - accuracy
update2=subset(update1, select = c("subjno","Group","Scores.spanEmo_T2_1","Scores.spanEmo_T2_2"))
update_long_acc=melt(update2, id = c("Group","subjno"),
                     measured = c("Scores.spanEmo_T2_1","Scores.spanEmo_T2_2"))

update_long_acc$time=factor(update_long_acc$variable, levels=c("Scores.spanEmo_T2_1","Scores.spanEmo_T2_2"),
                       labels = c("pre","post"))
names(update_long_acc)[names(update_long_acc)=="value"]="span_emo_acc"
update_long_acc$variable=NULL
update_long_acc=na.omit(update_long_acc)
summary(update_long_acc)

# mean accuracy by group and time
summaryBy(span_emo_acc ~ Group + time, FUN= mean, na.rm=TRUE, data=update_long_acc)
```

```{r}
# H2: span_emo_acc model 
m.2.update=lmer(span_emo_acc ~ Group*time + (1|subjno), update_long_acc)
summary(m.2.update) # Nothing significant
```
```{r}
mod=data.frame(emmeans(m.2.update, pairwise ~ time|Group, adjust="Bonferroni", type="response")$emmeans)
ggplot(update_long_acc, aes(Group, span_emo_acc, color=time, fill=time))+ ylim(0,10)+
  geom_violin(alpha=0.3)+
  geom_jitter(position=position_jitterdodge(dodge.width = 0.9), size=1, alpha=0.1)+
  geom_errorbar(data=mod, aes(Group, emmean, ymin = lower.CL, ymax = upper.CL),width = 0.05,size  = 0.5, position = position_dodge(width=0.9))+
  theme_bw()
```


### Set-shifting
```{r}
# Import data
shift=read.csv("~/Dropbox/PhD/04_phd-studies/Affective Control Training Study/ac_training/processed_data/setshift.csv")
shift$X=NULL
shift$subjno=as.factor(shift$subjno)

shift1=merge(demo,shift, by="subjno")
```

```{r}
# H2: shifting - accuracy
shift2=subset(shift1, select = c("subjno","Group","Scores.emoColNumRandErr_T5_1","Scores.emoColNumRandErr_T5_2"))
shift_long_acc=melt(shift2, id = c("Group","subjno"),
                     measured = c("Scores.emoColNumRandErr_T5_1","Scores.emoColNumRandErr_T5_2"))

shift_long_acc$time=factor(shift_long_acc$variable, levels=c("Scores.emoColNumRandErr_T5_1","Scores.emoColNumRandErr_T5_2"),
                       labels = c("pre","post"))
names(shift_long_acc)[names(shift_long_acc)=="value"]="RandErr_emo_acc"
shift_long_acc$variable=NULL
shift_long_acc=na.omit(shift_long_acc)
summary(shift_long_acc)

# mean accuracy by group and time
summaryBy(RandErr_emo_acc ~ Group + time, FUN= mean, na.rm=TRUE, data=shift_long_acc)
```

```{r}
# H2: RandErr_emo_acc model 
m.2.shift=lmer(RandErr_emo_acc ~ Group*time + (1|subjno), shift_long_acc)
Anova(m.2.shift, type=3)
summary(m.2.shift) # Significant main effect of time
```
```{r}
# H2: post-hoc
emmeans(m.2.shift, pairwise ~ time, adjust="Bonferroni")

mod=data.frame(emmeans(m.2.shift, pairwise ~ time|Group, adjust="Bonferroni", type="response")$emmeans)
ggplot(shift_long_acc, aes(Group, RandErr_emo_acc, color=time, fill=time))+
  geom_violin(alpha=0.3)+
  geom_jitter(position=position_jitterdodge(dodge.width = 0.9), size=1, alpha=0.1)+
  geom_errorbar(data=mod, aes(Group, emmean, ymin = lower.CL, ymax = upper.CL),width = 0.05,size  = 0.5, position = position_dodge(width=0.9))+
  theme_bw()


mod=data.frame(emmeans(m.2.shift, pairwise ~ time, adjust="Bonferroni")$emmeans)
ggplot(shift_long_acc, aes(time, RandErr_emo_acc, color=time, fill=time))+
  geom_violin(alpha=0.3)+
  geom_jitter(position=position_jitterdodge(), size=1, alpha=0.1)+
  geom_errorbar(data=mod, aes(time, emmean, ymin = lower.CL, ymax = upper.CL),width = 0.05,size  = 0.5, position = position_dodge(width=0.9))+
  theme_bw()
```
Significant reduction in random error post-training (pre-post: contrast=.05, SE=.02, p=.003).

```{r}
# H2a: shifting - RT
shift3=subset(shift1, select = c("subjno","Group","Scores.emoColNumRandErrRT_T5_1","Scores.emoColNumRandErrRT_T5_2"))
shift_long_RT=melt(shift3, id = c("Group","subjno"),
                     measured = c("Scores.emoColNumRandErrRT_T5_1","Scores.emoColNumRandErrRT_T5_2"))

shift_long_RT$time=factor(shift_long_RT$variable, levels=c("Scores.emoColNumRandErrRT_T5_1","Scores.emoColNumRandErrRT_T5_2"),
                       labels = c("pre","post"))
names(shift_long_RT)[names(shift_long_RT)=="value"]="RandErr_emo_RT"
shift_long_RT$variable=NULL
shift_long_RT=na.omit(shift_long_RT)
summary(shift_long_RT)

# mean accuracy by group and time
summaryBy(RandErr_emo_RT ~ Group + time, FUN= mean, na.rm=TRUE, data=shift_long_RT)
```

```{r}
# H2a: RanErr_emo_RT model
m.2a.shift=lmer(RandErr_emo_RT ~ Group*time + (1|subjno),shift_long_RT)
Anova(m.2a.shift, type=3)
summary(m.2a.shift) # significant main effect of time (p >.001) and marginal interaction (p=.069)
```

```{r}
# H2a: post-hoc
emmeans(m.2a.shift, pairwise ~ time, adjust="Bonferroni")
emmeans(m.2a.shift, pairwise ~ time|Group, adjust="Bonferroni")

mod=data.frame(emmeans(m.2a.shift, pairwise ~ time|Group, adjust="Bonferroni", type="response")$emmeans)
ggplot(shift_long_RT, aes(Group, RandErr_emo_RT, color=time, fill=time))+
  geom_violin(alpha=0.3)+
  geom_jitter(position=position_jitterdodge(dodge.width = 0.9), size=1, alpha=0.1)+
  geom_errorbar(data=mod, aes(Group, emmean, ymin = lower.CL, ymax = upper.CL),width = 0.05,size  = 0.5, position = position_dodge(width=0.9))+
  theme_bw()


mod=data.frame(emmeans(m.2a.shift, pairwise ~ time, adjust="Bonferroni")$emmeans)
ggplot(shift_long_RT, aes(time, RandErr_emo_RT, color=time, fill=time))+
  geom_violin(alpha=0.3)+
  geom_jitter(position=position_jitterdodge(), size=1, alpha=0.1)+
  geom_errorbar(data=mod, aes(time, emmean, ymin = lower.CL, ymax = upper.CL),width = 0.05,size  = 0.5, position = position_dodge(width=0.9))+
  theme_bw()
```

*** 
## H3: Age-related change hypothesis
Adding age as a moderator in the H1 model. 
```{r}
m.3= lmer(dPrimeEmo ~ Group*time*decimal_age + (1|subjno), 
          data=nback_long)
Anova(m.3, type=3)
summary(m.3)
```

```{r}
# Age as a polynomial 
m.3a=lmer(dPrimeEmo ~ Group*time*poly(decimal_age,3) + (1|subjno),
          data=nback_long)
summary(m.3a)
```


```{r}
# H3: post-hocs
# Main effect of time
emmeans(m.3, pairwise~time, adjust="Bonferroni")

# group x time 
emmeans(m.3, pairwise~time|Group, adjust="Bonferroni") #significant interactions in the model but none of the contrasts is significant - what does this mean? 
emmeans(m.3, pairwise~Group|time, adjust="Bonferroni") #no difference between groups at both time points 

table = summaryBy(dPrimeEmo ~ time+Group, nback_long, FUN=mean)
ggplot(nback_long, aes(time, dPrimeEmo, color=time, fill=time))+facet_wrap(~Group)+
  geom_boxplot(alpha=0.2)+geom_smooth(method='lm')+theme_bw()
```


```{r}
# age x time
emmeans(m.3,pairwise~time|decimal_age, adjust="Bonferroni")

table=summaryBy(dPrimeEmo ~ decimal_age+time, nback_long, FUN=mean)
ggplot(table, aes(decimal_age, dPrimeEmo.mean, color=time, fill=time))+
  geom_point(alpha=0.2)+geom_smooth(method='lm')+theme_bw()
```


```{r}
# time x age x group 
table <- summaryBy(dPrimeEmo ~ decimal_age + time + Group, nback_long, FUN=mean)
ggplot(table, aes(decimal_age, dPrimeEmo.mean, color=time, fill=time))+facet_wrap(~Group)+
  geom_point(alpha=0.2) + geom_smooth(method = 'lm')+theme_bw()
```

## H4: Mental health hypothesis
Increase in affective control from pre- to post-training will be associated with fewer MH problems and emotion regulation difficulties, as well as higher levels of self-reported self-control. 

```{r}
# Import questionnaire data 
t1=read.csv("~/Dropbox/PhD/04_phd-studies/Affective Control Training Study/ac_training/processed_data/t1_questionnaire_scored.csv")
t2=read.csv("~/Dropbox/PhD/04_phd-studies/Affective Control Training Study/ac_training/processed_data/t2_questionnaire_scored.csv")

t1$X=NULL
t2$X=NULL
```

```{r}
# SDQ - tidying
sdq_t1=subset(t1, select=c("subjno","sdq_sum_t1"))
sdq_t2=subset(t2, select=c("subjno","sdq_sum_t2"))
sdq=merge(sdq_t1,sdq_t2, by="subjno")

sdq_long=melt(sdq, id = c("subjno"),
              measured = c("sdq_sum_t1","sdq_sum_t2"))  
sdq_long$time=factor(sdq_long$variable, levels = c("sdq_sum_t1","sdq_sum_t2"),
                     labels = c("pre","post"))
names(sdq_long)[names(sdq_long)=="value"]="sdq_sum"
sdq_long$variable=NULL
sdq_long=na.omit(sdq_long)
summary(sdq_long)
```

```{r}
nback_sdq=merge(nback_long, sdq_long, by=c("subjno","time"))
summary(nback_sdq)
```

```{r}
mh.1=lmer(sdq_sum~dPrimeEmo + time + decimal_age + (1|subjno),
          data=nback_sdq)
summary(mh.1)
```






