---
title: "Affective Control Training: Analyses from Protocol"
author: "Jovita T. Leung"
date: "28/03/2021"
output: 
  html_document: 
  toc: yes
  toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup
```{r message=F, warning=F}
rm(list=ls())

# Load libraries
library(dplyr)
library(doBy)
library(lmerTest) 
library(Hmisc)
library(anchors)
library(car) # Anova
library(emmeans)
library(ggfortify)
library(data.table)
library(psych)
library(ggeffects) # extract model predictions
library(sjPlot) # make LMM summary tables
library(jtools) # summarise LMM models
library(ggsignif)
library(ggpubr) # arrange plots
library(Bolstad) # bayesian t test
library(reshape2)
```

```{r}
# Import data
demo=read.csv("~/Dropbox/PhD/Affective Control Training Study/ac_training/processed_data/demographics_group.csv")
nback=read.csv("~/Dropbox/PhD/Affective Control Training Study/ac_training/processed_data/nback.csv")
train=read.csv("~/Dropbox/PhD/Affective Control Training Study/ac_training/processed_data/training_info.csv")
all=read.csv("~/Dropbox/PhD/Affective Control Training Study/ac_training/data/281019wodupl2.csv") # full data from Susanne
```


```{r warning=F, message=F}
demo$X=NULL
demo[c("subjno","gender","ses","Group")]=lapply(demo[c("subjno","gender","ses","Group")],factor)

train$X=NULL
train[c("subjno","no_of_sessions_under_10_mins","no_of_sessions_over_10_mins","no_of_moodtracker_only_sessions",
        "total_number_of_days_1_session_only","total_number_of_days_2plus_sessions","days_betw_T1T2")] = lapply(
          train[c("subjno","no_of_sessions_under_10_mins","no_of_sessions_over_10_mins","no_of_moodtracker_only_sessions",
                  "total_number_of_days_1_session_only","total_number_of_days_2plus_sessions","days_betw_T1T2")], 
          factor)

nback$X=NULL
nback$subjno=as.factor(nback$subjno)
nback[2:37]=lapply(nback[2:37],as.numeric)
```

```{r}
nback1=merge(demo,nback, by="subjno")
```


## Training groups comparison
Following the protocol, two training groups will be compared for potential confounds before hypothesis testing. Any variables showing significant group differences at baseline will be added to subsequent group comparison as covariate. 

```{r}
# Age
bayes.t.test(decimal_age ~ Group, demo) #no significant group difference (p=.202)
```

```{r}
# Gender
chisq.test(demo$gender,demo$Group) #no significant group difference (p=.616)
```

```{r}
# Parental education
plot(demo$ses)
chisq.test(demo$ses, demo$Group) #no significant group difference (p=.902)
```

```{r}
# IQ
plot(demo$iq_sum)
summary(glm(demo$iq_sum~demo$Group)) #no significant group difference (p=.135)
```

```{r}
plot(demo$test_days)
summary(glm(demo$test_days~demo$Group)) #no significant group difference (p=.360)
```

## H1: Affective control training hypothesis
Affective control can be improved in adolescents. To investigate this hypothesis, we will compare individuals’ performances on the affective n-back task across the two training groups.

```{r}
# H1: nback - d prime 
nback2=subset(nback1, select=c("subjno","Group","Scores.dPrimeEmo_T4_1","Scores.dPrimeEmo_T4_2"))
nback_long<-reshape2::melt(nback2, id = c("Group", "subjno"),
                                 measured = c("Scores.dPrimeEmo_T4_1",
                                              "Scores.dPrimeEmo_T4_2"))

nback_long$time=factor(nback_long$variable, levels=c("Scores.dPrimeEmo_T4_1","Scores.dPrimeEmo_T4_2"),
                       labels = c("pre","post"))
names(nback_long)[names(nback_long)=="value"]="dPrimeEmo"
nback_long$variable=NULL
summary(nback_long)

# mean d' by group and time
summaryBy(dPrimeEmo ~ Group + time, FUN=mean, na.rm=TRUE, data=nback_long)
```
```{r}
# H1: dPrimeEmo model
m.1= lmer(dPrimeEmo ~ Group*time + (1|subjno), 
          data=nback_long)
summary(m.1) # nothing significant
```

```{r}
# H1a: nback - RT
nback3=subset(nback1, select=c("subjno","Group","Scores.meanRTCorrHitsEmo_T4_1","Scores.meanRTCorrHitsEmo_T4_2"))
nback_long_RT<-reshape2::melt(nback3, id = c("Group", "subjno"),
                                 measured = c("Scores.meanRTCorrHitsEmo_T4_1",
                                              "Scores.meanRTCorrHitsEmo_T4_2"))

nback_long_RT$time=factor(nback_long_RT$variable, levels=c("Scores.meanRTCorrHitsEmo_T4_1","Scores.meanRTCorrHitsEmo_T4_2"),
                       labels = c("pre","post"))
names(nback_long_RT)[names(nback_long_RT)=="value"]="meanRT_emo"
nback_long_RT$variable=NULL
summary(nback_long_RT)

# mean RT by group and time
summaryBy(meanRT_emo ~ Group + time, FUN=mean, na.rm=TRUE, data=nback_long_RT)
```

```{r}
# H1a: meanRT_emo model
m.1a= lmer(meanRT_emo ~ Group*time + (1|subjno), 
          data=nback_long_RT)
summary(m.1a)
```
```{r}
# H1a: post-hoc
emmeans(m.1a, pairwise~time, adjust="Bonferroni") #meanRT is significantly lower post-training
emmeans(m.1a, pairwise~time|Group, adjust="Bonferroni") #meanRT for both groups are lower post-training
```
Significant main effects of Time but not Group. Significant interactions between Group and Time. Mean RT is lower for both groups post-training - this is probably due to familiarity with the task so not very interesting. 


*** 
## H2: Affective control facets hypothesis
AC training will improve inhibition, updating and shifting. 

### Inhibition
```{r}
# Import data
inhib=read.csv("~/Dropbox/PhD/Affective Control Training Study/ac_training/processed_data/stroop.csv")
inhib$X=NULL
inhib$subjno=as.factor(inhib$subjno)

inhib1=merge(demo,inhib, by="subjno")
```

```{r}
# H2: Inhibition - accuracy
inhib2=subset(inhib1, select = c("subjno","Group","Scores.accSadIncongRaw_T3_1","Scores.accSadIncongRaw_T3_2"))
inhib_long_acc=melt(inhib2, id = c("Group","subjno"),
                     measured = c("Scores.accSadIncongRaw_T3_1","Scores.accSadIncongRaw_T3_2"))

inhib_long_acc$time=factor(inhib_long_acc$variable, levels=c("Scores.accSadIncongRaw_T3_1","Scores.accSadIncongRaw_T3_2"),
                       labels = c("pre","post"))
names(inhib_long_acc)[names(inhib_long_acc)=="value"]="acc_sad_incong"
inhib_long_acc$variable=NULL
summary(inhib_long_acc)

# mean accuracy by group and time
summaryBy(acc_sad_incong ~ Group + time, FUN= mean, na.rm=TRUE, data=inhib_long_acc)
```

```{r}
# H2: acc_sad_incong model 
m.2=lmer(acc_sad_incong ~ Group*time + (1|subjno), inhib_long_acc)
summary(m.2) # Nothing significant
```

```{r}
# H2a: Inhibition - RT
inhib3=subset(inhib1, select = c("subjno","Group","Scores.RTSadIncong_T3_1","Scores.RTSadIncong_T3_2"))
inhib_long_RT=melt(inhib3, id = c("Group","subjno"),
                     measured = c("Scores.RTSadIncong_T3_1","Scores.RTSadIncong_T3_2"))

inhib_long_RT$time=factor(inhib_long_RT$variable, levels=c("Scores.RTSadIncong_T3_1","Scores.RTSadIncong_T3_2"),
                       labels = c("pre","post"))
names(inhib_long_RT)[names(inhib_long_RT)=="value"]="RT_sad_incong"
inhib_long_RT$variable=NULL
summary(inhib_long_RT)

# mean accuracy by group and time
summaryBy(RT_sad_incong ~ Group + time, FUN= mean, na.rm=TRUE, data=inhib_long_RT)
```

```{r}
# H2a: RT_sad_incong model
m.2a=lmer(RT_sad_incong ~ Group*time + (1|subjno), inhib_long_RT)
summary(m.2a)
```

```{r}
# H2a: post-hoc
emmeans(m.2a, pairwise~time, adjust="Bonferroni")
```
There is a significant main effect of time on sad_incong RT, where RT is lower post-training. No signification interaction with group and time. 

### Updating
```{r}
# Import data
update=read.csv("~/Dropbox/PhD/Affective Control Training Study/ac_training/processed_data/digit_span.csv")
update$X=NULL
update$subjno=as.factor(update$subjno)

update1=merge(demo,update, by="subjno")
```

```{r}
# H2: updating - accuracy
update2=subset(update1, select = c("subjno","Group","Scores.spanEmo_T2_1","Scores.spanEmo_T2_2"))
update_long_acc=melt(update2, id = c("Group","subjno"),
                     measured = c("Scores.spanEmo_T2_1","Scores.spanEmo_T2_2"))

update_long_acc$time=factor(update_long_acc$variable, levels=c("Scores.spanEmo_T2_1","Scores.spanEmo_T2_2"),
                       labels = c("pre","post"))
names(update_long_acc)[names(update_long_acc)=="value"]="span_emo_acc"
update_long_acc$variable=NULL
summary(update_long_acc)

# mean accuracy by group and time
summaryBy(span_emo_acc ~ Group + time, FUN= mean, na.rm=TRUE, data=update_long_acc)
```

```{r}
# H2: span_emo_acc model 
m.2.update=lmer(span_emo_acc ~ Group*time + (1|subjno), update_long_acc)
summary(m.2.update) # Nothing significant
```

### Set-shifting
```{r}
# Import data
shift=read.csv("~/Dropbox/PhD/Affective Control Training Study/ac_training/processed_data/setshift.csv")
shift$X=NULL
shift$subjno=as.factor(shift$subjno)

shift1=merge(demo,shift, by="subjno")
```

```{r}
# H2: shifting - accuracy
shift2=subset(shift1, select = c("subjno","Group","Scores.emoColNumRandErr_T5_1","Scores.emoColNumRandErr_T5_2"))
update_long_acc=melt(update2, id = c("Group","subjno"),
                     measured = c("Scores.emoColNumRandErr_T5_1","Scores.emoColNumRandErr_T5_2"))

update_long_acc$time=factor(update_long_acc$variable, levels=c("Scores.emoColNumRandErr_T5_1","Scores.emoColNumRandErr_T5_2"),
                       labels = c("pre","post"))
names(update_long_acc)[names(update_long_acc)=="value"]="span_emo_acc"
update_long_acc$variable=NULL
summary(update_long_acc)

# mean accuracy by group and time
summaryBy(span_emo_acc ~ Group + time, FUN= mean, na.rm=TRUE, data=update_long_acc)
```

```{r}
# H2: span_emo_acc model 
m.2.update=lmer(span_emo_acc ~ Group*time + (1|subjno), update_long_acc)
summary(m.2.update) # Nothing significant
```



